### 2024-12-03 paper/paper.qmd
Initial setup to transfer over local work

---
title: ""
author: 
  - Jerry Xia
thanks: "Code and data are available at: <>"
date: today
date-format: long
abstract: ""
format: pdf
number-sections: true
toc: true 
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
```